Visually pleasing knot projections

### 2022-06-07 inst/jma_rebuttal_review3.tex

Page 4 line 48: ``However, if we can ensure that the radius of
curvature changes only by a small amount,..."  -$>>$ HOW can the user do
this?

Page 4 lines 50-52: No need to differentiate here between over- and
under-strands.  I assume they are treated the same in the optimization
process.

Page 4 lines 53-54: ``We would like strand crossing points to be far
from nodes,..."  -- Why?  The curves are also visually continuous across
the nodes!

\textcolor{blue}{The curves certainly are visually continuous across
  the nodes---but only in the optimized knots.  The motivation for
  keeping Bezier nodes far from crossing points is twofold.  Firstly,
  as one traces a curve visually, Bezier nodes cannot possibly avoid
  introducing {\em some} form of disjointness in the path; and such
  disjointness is particularly disruptive if on the understrand of a
  node, as the visual continuity is interrupted.  I have added a
  figure (including two different diagrams of a trefoil knot) to the
  Gallery that illustrates the reason why nodes' being at crossing
  points is undesirable.  In short, at the node of a Bezier curve, the
  radius of curvature undergoes a visually jarring discontinuity which
  looks bad, especially the understrand, which is particularly
  susecptible to such considerations because of the break.  I do
  appreciate that many viewers will find the distinction invisible or
  inconsequential; but my informal polling of family and colleagues
  reveals that, once sensitised to the issue, they cannot unsee the
  kink.  \\ \\ The second reason is more mundane.  With reference to
  figure 7 (the four figure-of-eight knots diagram), we can see a
  vertical line of symmetry on which there are two crossing points.
  Imagine there was a node at one of these crossing points.  There
  would have to be {\em two} nodes at the same point, one for the
  overstrand and one for the understrand; the software would somehow
  have to impose vertical mirror symmetry on the diagram and this
  would necessitate that the knot class contain yet another symmetry
  object, alongside the 12 existing ones.  These considerations become
  yet more onerous when considering rotational symmetry as in Figure 8
  (which shows $5_1$).  If there were a node at, say, the intersection
  of strand 18-19 and 9-10 then somehow the two nodes would have to
  embody not only mirror symmetry (with respect to 5-6/14-15) but also
  fivefold rotational symmetry.  Simply put, this is a nightmare.  The
  software documentation contains a substantial discussion of the
  issue, but I do not believe that the manuscript would benefit from
  what amounts to little more than me grumbling about implementation
  difficulties.}

### 2022-06-07 inst/jma_knot.tex
fixes #58

Here, knot paths are required to have Bezier handles that are
symmetrically placed with regard to nodes (this is a user-settable
option in inkscape).  One consequence of this design choice is that
radius of curvature is not matched exactly across a node; the path is
not G2 smooth in CAD terminology.  However, if we can ensure that the
radius of curvature changes only by a small amount, visual
continuity---in the sense of absence of visible interruptions to a
smoothly evolving strand---is preserved.  At a crossing point it is
important that both the unbroken and broken strands (designated
``overstrand'' and ``understrand'' respectively) are smooth.  We would
like strand crossing points to be far from nodes, as this ensures
visual continuity of strands at crossing points, especially the
understrand (see Figure~\ref{trefoil_circles} for an example and
discussion).

\begin{figure}[h]
  \centering
    \includegraphics[width=10cm]{trefoil_circles_post_inkscape.pdf}
\caption{A trefoil knot \label{trefoil_circles} constructed from arcs
  of circles (elastica) constrained to cross at right angles; the
  strands exert a couple but no force on one another.  Mirror symmetry
  and threefold rotational symmetry are imposed.  Elementary
  trigonometry shows that the radii of the inner and outer loops are
  $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$ and
  $\sec(\pi/12)=\sqrt{6}-\sqrt{2}$ respectively.  The curvature thus
  has a displeasing discontinuity at a node which interrupts the
  visual continuity there, especially the understrand}
\end{figure}

\begin{figure}[h]
  \centering  
  % To create file "two_trefoil_knots_post_inkscape.pdf", run
  % knotter.R, which makes (among other things) two_trefoil_knots.pdf.
  % Edit this in inkscape [make a mirror image, then change the
  % colours and finally use the raise and lower tools to put
  % everything in the right order]
    \includegraphics[width=10cm]{two_trefoil_knots_post_inkscape.pdf}
    \caption{Two trefoil knots \label{two_trefoil_knots}.  Left,
      optimized as per the {\tt knotR} package techniques documented
      here; right, constructed from arcs of circles constrained to
      cross at right angles.  The strands exert a couple but no force
      on one another.  Mirror symmetry and threefold rotational
      symmetry are imposed.  Elementary trigonometry shows that the
      radii of the inner and outer loops are
      $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$ and
      $\sec(\pi/12)=\sqrt{6}-\sqrt{2}$ respectively.  The curvature
      thus has a displeasing discontinuity at a node which interrupts
      the visual continuity there, especially the understrand}
\end{figure}

@@

There now follows a selection of pleasing knot diagrams taken from
datasets provided with the package, illustrating the relative ease
with which knot diagrams may be created and optimized.
Figure~\ref{perko_AB} shows two forms of a 10-crossing knot once
erroneously considered to be topologically distinct;
figure~\ref{orn20} shows an ornamental knot with exact five-fold
symmetry (but not dihedral symmetry); and figure~\ref{all8} shows all
prime knots with eight or fewer crossings, following
Rolfsen~\cite{rolfsen1976}.  The package includes all knots to nine
crossings.

>>>>

Here, knot paths are required to have Bezier handles that are
symmetrically placed with regard to nodes (this is a user-settable
option in inkscape).  One consequence of this design choice is that
radius of curvature is not matched exactly across a node; the path is
not G2 smooth in CAD terminology.  However, if we can ensure that the
radius of curvature changes only by a small amount, visual
continuity---in the sense of absence of visible interruptions to a
smoothly evolving strand---is preserved.  At a crossing point it is
important that both the unbroken and broken strands (designated
``overstrand'' and ``understrand'' respectively) are smooth.  We would
like strand crossing points to be far from nodes, as this ensures
visual continuity of strands at crossing points, especially the
understrand (see Figure~\ref{two_trefoil_knots} for an example and
discussion).

@@

There now follows a selection of pleasing knot diagrams taken from
datasets provided with the package, illustrating the relative ease
with which knot diagrams may be created and optimized.
Figure~\ref{two_trefoil_knots} shows and compares two plausible
schemes for rendering a trefoil.  Figure~\ref{perko_AB} shows two
forms of a 10-crossing knot once erroneously considered to be
topologically distinct; figure~\ref{orn20} shows an ornamental knot
with exact five-fold symmetry (but not dihedral symmetry); and
figure~\ref{all8} shows all prime knots with eight or fewer crossings,
following Rolfsen~\cite{rolfsen1976}.  The package includes all knots
to nine crossings.

\begin{figure}[h]
  \centering  
  % To create file "two_trefoil_knots_post_inkscape.pdf", run
  % knotter.R, which makes (among other things) two_trefoil_knots.pdf.
  % Edit this in inkscape [make a mirror image, then change the
  % colours and finally use the raise and lower tools to put
  % everything in the right order]
    \includegraphics[width=10cm]{two_trefoil_knots_post_inkscape.pdf}
    \caption{Two trefoil \label{two_trefoil_knots} knots.  Left,
      optimized as per the {\tt knotR} package techniques documented
      here; right, constructed from arcs of circles constrained to
      cross at right angles.  The strands exert a couple but no force
      on one another.  Mirror symmetry and threefold rotational
      symmetry are imposed.  Elementary trigonometry shows that the
      radii of the inner and outer loops are
      $\operatorname{cosec}(\pi/12)=\sqrt{6}+\sqrt{2}$ and
      $\sec(\pi/12)=\sqrt{6}-\sqrt{2}$ respectively.  The curvature
      thus has a displeasing discontinuity at a node which interrupts
      the visual continuity there, especially the understrand}
\end{figure}
